---
title: "Course Project 1 - Reproducible Research"
author: "Mike"
date: "11/16/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r getdata, echo=TRUE}
# Set working directory for file directory
# Load the data (i.e. read.csv(), read.csv())
fulldata<-read.table("activity.csv", header=TRUE, sep=",", na.strings="NA")
```

# Process/transform the data (if necessary) into a format suitable for your analysis
```{r cleandata, echo=TRUE}
cleandata<-fulldata
cleandata$interval<-as.factor(cleandata$interval)
cleandata$steps<-as.numeric(cleandata$steps)
cleandata<-cleandata[complete.cases(cleandata),]
```

# Calculate the total number of steps taken per day
```{r TotalStepsPerDay, echo=TRUE}
library(dplyr)
stepsxday<-cleandata%>%group_by(date)%>%summarize(sum(steps))

# Make a histogram of the total number of steps taken each day
hist(stepsxday$`sum(steps)`,main="Histogram of Total # of Steps Each Day", xlab="Sum Steps")

# Calculate and report the mean and median of the total number of steps taken per day
summary(stepsxday$`sum(steps)`)
```

# Make a time series plot (i.e.,type = "l") of the 5-minute interval (x-axis)...
# and the average number of steps taken, averaged across all days (y-axis)
```{r StepsPerIntervals, echo=TRUE}
stepsxinterval<-cleandata%>%group_by(interval)%>%summarize(mean(steps))
plot(stepsxinterval$interval, stepsxinterval$`mean(steps)`, type="l", xlab="avg steps", ylab="interval")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
maxinterval<-which.max(stepsxinterval$`mean(steps)`)
stepsxinterval[maxinterval,]
```

# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r NAvalues, echo=TRUE}
sum(!complete.cases(fulldata))

# Devise a strategy for filling in all of the missing values in the dataset.
# (replace NA values with '0')
# Create a new dataset that is equal to the original dataset but with the missing data filled in.
newdata<-fulldata
newdata[is.na(newdata)]<-0
```

# Make a histogram of the total number of steps taken each day... 
```{r newStepsPerDay, echo=TRUE}
newstepsxday<-newdata%>%group_by(date)%>%summarize(sum(steps))
hist(newstepsxday$`sum(steps)`,main="Histogram of Total # of Steps Each Day", xlab="Sum Steps")

# Calculate and report the mean and median total number of steps taken per day (for new data)...
summary(newstepsxday$`sum(steps)`)
```

# Do these values differ from the estimates from the first part of the assignment?
# ANSWER:  YES

# What is the impact of imputing missing data on the estimates of the total daily number of steps?
# ANSWER:  Replacing NA values with zero: decreased the minimum value, mean and median; increased the range.

# Are there differences in activity patterns between weekdays and weekends?
# Create a new factor variable in the dataset with two levels - "weekday" and "weekend"...
# indicating whether a given date is a weekday or weekend day.
```{r weekfactor, echo=TRUE}
weekdata<-newdata
weekdata$date<-as.Date(weekdata$date)
weekdata$date<-weekdays(weekdata$date)
weekdata$date<-as.character(weekdata$date)

weekdata <- mutate(weekdata, weekCAT=weekdata$date)
weekdata$weekCAT[weekdata$weekCAT=="Saturday"|weekdata$weekCAT=="Sunday"]<-"weekend"
weekdata$weekCAT[weekdata$weekCAT!="weekend"]<-"weekday"
weekdata$interval<-as.factor(weekdata$interval)
```

#weekdata$weekCAT<-as.factor(weekdata$weekCAT)

# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and...
# the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r weekCAT, echo=TRUE}
weekENDsteps<-subset(weekdata, weekCAT=="weekend")
weekDAYsteps<-subset(weekdata, weekCAT=="weekday")

stepsxEND<-weekENDsteps%>%group_by(interval)%>%summarize(mean(steps))
stepsxDAY<-weekDAYsteps%>%group_by(interval)%>%summarize(mean(steps))

par(mfrow=c(2,1))
plot(stepsxEND$interval, stepsxEND$`mean(steps)`, type="l", xlab="avg steps", ylab="interval",main="weekend")
plot(stepsxDAY$interval, stepsxDAY$`mean(steps)`, type="l", xlab="avg steps", ylab="interval",main="weekday")
```
